---
title: 'Primary school attendance rates of children by their disability status: <br> Insights observed from data of Myanmar Census(2014)'
author: "UNICEF Myanmar, SPCRM"
date: "2023-06-07"
always_allow_html: true
output: github_document
---

# Overview

This analysis is based on the *township level* information data of Myanmar Census 2014.We are interested in examining the difference of school attendance rates between the children with *no disability* and their counterparts with *severe disability*.

Using the **REDATAM** that is designed to extract data from Census, we constructed two data sets at township level that are representing the primary school attendance rates of our groups of interest— normal vs. severe disability of the primary school age population of the 6-10 year old children.

State/Region aggregates are calculated.The results are visualized for comparing the situations of the groups.

All datasets produced in our analytical project can be found on the Github repository.We also provided a couple of R-script documents that we used during the process for data arrangement and analyses. *https://github.com/AungLwin25/Disability*

Particular credits to our colleagues, Bjorn Gelders for formulating the framework of analysis, and Aung Lwin for kniting algorithms, wrangling data, computing statistics, and producing visuals.
                                          — UNICEF Myanmar, SPCRM

```{r setup, include=FALSE, echo=FALSE}
library(readxl)
library(haven)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(janitor)
library(ggthemes)
library(gridExtra)
library(reshape2)

knitr::opts_chunk$set(echo = TRUE)
```

```{r read, echo=FALSE, message=FALSE, results='hide',warning=FALSE}
# Gather base township information (geomarkers, districts,states/regions)
TsDB128b <- read_dta("C:/Users/aklwin/OneDrive/UNICEF ME2/Database/States/Profile/Townships/Data/Work/TsDB128b.dta")

Ts<-TsDB128b

library(tidyverse)
Ts %>% select(5,9,10,11,13,14) -> Loc

rename(Loc, SR=MyaState2,
             AreaCode = TsAreaCode) -> Loc2

```

```{r mutate_var, echo=FALSE,message=FALSE, results='hide',warning=FALSE}
# Gather disability data set and normal data set
## Disability
TsSex <- function(name) {
    temp <<- readxl::read_excel(file.path(directory = "C:/Users/aklwin/OneDrive/R programming/Github/Disability", paste0(name,".xlsx")))
    temp[1:413,1:5] ->> temp2
    write.csv(temp2, paste0(name,"_cl",".csv"))
    temp2 ->temp3
}

TsSex("Cs14AnySevereDsbPriSex") -> df
TsSex("Cs14AnySevereDsbPriAttSex") -> df2

df[1] -> a
df2[1] -> b
sum(a==b)

rename(df, SevMale = Male,
           SevFemale= Female,
           SevTotal = Total) -> d1

df2 %>% select(3:5) %>% rename_with( ~ paste0("AttSv", .x)) -> d2

cbind(d1,d2) -> d3
d3 %>% mutate( pcAttSvMale   = 100 * AttSvMale/SevMale,
               pcAttSvFemale = 100 * AttSvFemale/SevFemale,
               pcAttSvTotal = 100 * AttSvTotal/SevTotal) -> d4

## Normal
TsSex("Cs14_NoDsbPriAge") -> n1
TsSex("Cs14NoDsbPriAttSex") -> n2

n1 %>% select(3:5) %>% rename_with( ~ paste0("Nm", .x)) -> n3
n2 %>% select(3:5) %>% rename_with( ~ paste0("AttNm", .x)) -> n4

cbind(n1[1:2],n3,n4) -> n5
head(n5,3)

n5 %>% mutate( pcAttNmMale   = 100 * AttNmMale/NmMale,
               pcAttNmFemale = 100 * AttNmFemale/NmFemale,                
               pcAttNmTotal = 100 * AttNmTotal/NmTotal) -> n6

# Combine two data sets
cbind(d4,n6[3:11]) -> PriAtt

rename(PriAtt,AreaCode = `Area code`)-> p2
inner_join(Loc2,p2, by = "AreaCode") -> p3
```

# Average primary school attendances of townships 
## Tables

### Summary statistics
```{r sum, echo=FALSE}
p3 %>% group_by(SR) %>%
summarise(TsAvgNmPri= mean(pcAttNmTotal,na.rm=TRUE),
TsAvgSvPri= mean(pcAttSvTotal,na.rm=TRUE)) -> pcSRts_pri

rename(pcSRts_pri,Normal = TsAvgNmPri,
                 Severe = TsAvgSvPri) ->Pri3

summary(Pri3)
```

### State/region aggregates
```{r table, echo=FALSE}
# Summarize state/region aggregates

print(Pri3)
```

## Visuals
### Boxplot
```{r bp, echo=FALSE, message=FALSE, results='hide',warning=FALSE}
melt(pcSRts_pri, id="SR") -> pcSRts_pri2

pcSRts_pri2$v2 <-ifelse(pcSRts_pri2$variable=="TsAvgNmPri", "Normal","Severe")
```

```{r bp2, echo=FALSE}
ggplot(pcSRts_pri2, aes(y=value, fill=v2)) +
    geom_boxplot(width=0.5) +
    labs(fill="Disability status") +
    scale_x_discrete(labels = NULL, breaks = NULL) +
    ylab("percent") +
    theme_minimal() +
    theme(legend.position = c(0.2, 0.5)) 
```

### Group bar
```{r gb, echo=FALSE,message=FALSE, results='hide',warning=FALSE}
melt(pcSRts_pri, id="SR") -> pcSRts_pri2
pcSRts_pri2
pcSRts_pri2$v2 <-ifelse(pcSRts_pri2$variable=="TsAvgNmPri", "Normal","Severe")
```

```{r gb2, echo=FALSE}
ggplot(pcSRts_pri2, aes(y=value, x=SR, fill = v2)) +
    geom_bar(position="dodge", stat="identity") +
    scale_fill_manual(name="",
                      values=c("#99CCFF","#FF9999"),
                      labels=c("No disability","Severe disability"))  +
    xlab ("State and Region") + ylab ("Percent") + 
    theme_minimal()+
    theme(legend.position = "top") + 
    theme(axis.text.x = element_text(angle=90)) +
    ggtitle("Primary School Attendance")
```

# Conclusion
The analytical results from Myanmar Census 2014 show that the children having *severe disability* were **hugely** lagging behind in their primary school attendance compared to those of the children with *no disability*.

# Appendix

## Data Table of township values
```{r print_tb, echo=FALSE, results= 'asis', warning=FALSE, message=FALSE}
p3 %>% select(1,4,contains("pc")) -> p4
names(p4) <-c("State/Region","Census Township", "Male Severe","Female Severe","Total Severe","Male Normal","Female Normal","Total Normal")
```
```{r print_tb2, echo=FALSE, results= 'asis', warning=FALSE, message=FALSE}
options(digits = 3)
kable(p4, caption= "<Heading 1> Primary School Attendance rates",format ="html") %>%  
    footnote(general = "", general_title = "N.B. The NA values are on the children with disaiblity side that might also mean their being neglected status.")
```









