---
title: 'School attendance rates of children in primary school age by their disability status: <br> Insights observed from data of Myanmar Census(2014)'
author: "UNICEF Myanmar, SPCRM"
date: "2023-06-08"
always_allow_html: true
output: github_document
---

# Overview

This analysis is based on the *township level* information data of Myanmar Census 2014.We are interested in examining the difference of school attendance rates between the children with *no disability* and their counterparts with *severe disability* for the population of primary school age children (6-10 year old).

Using the **REDATAM** that is designed to extract data from Census, we constructed two data sets at township level that are representing the primary school attendance rates of our targeted groups of analysis — normal vs. severe disability of the primary school age population.

The State/Region aggregates are calculated in two ways.The first method finds the average attendance rate of townships of state/regions in each group. The second method directly computes each group's state/regions' attendance rates from their total counts of attending children divided by the respective base populations.In the current post for web display, we focused on the results of the *weighted analysis* that is applicable with the second method.The results are visualized for comparing the situations of the groups.

All data sets produced in our analytic project can be found on the Github repository.We also provided a couple of R-script documents that we used during the process for data arrangement and analyses. *https://github.com/AungLwin25/Disability*

Particular credits to our colleagues, Bjorn Gelders for formulating the framework of analysis, and Aung Lwin for kniting algorithms, wrangling data, computing statistics, and producing visuals.
                                          — UNICEF Myanmar, SPCRM

```{r setup, include=FALSE, echo=FALSE}
library(readxl)
library(haven)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(janitor)
library(ggthemes)
library(gridExtra)
library(reshape2)

knitr::opts_chunk$set(echo = TRUE)
```

```{r read, echo=FALSE, message=FALSE, results='hide',warning=FALSE}
# Gather base township information (geomarkers, districts,states/regions)
TsDB128b <- read_dta("C:/Users/aklwin/OneDrive/UNICEF ME2/Database/States/Profile/Townships/Data/Work/TsDB128b.dta")

Ts<-TsDB128b

library(tidyverse)
Ts %>% select(5,9,10,11,13,14) -> Loc

rename(Loc, SR=MyaState2,
             AreaCode = TsAreaCode) -> Loc2

```

```{r mutate_var, echo=FALSE,message=FALSE, results='hide',warning=FALSE}
# Gather disability data set and normal data set
## Disability
TsSex <- function(name) {
    temp <<- readxl::read_excel(file.path(directory = "C:/Users/aklwin/OneDrive/R programming/Github/Disability", paste0(name,".xlsx")))
    temp[1:413,1:5] ->> temp2
    write.csv(temp2, paste0(name,"_cl",".csv"))
    temp2 ->temp3
}

TsSex("Cs14AnySevereDsbPriSex") -> df
TsSex("Cs14AnySevereDsbPriAttSex") -> df2

df[1] -> a
df2[1] -> b
sum(a==b)

rename(df, SevMale = Male,
           SevFemale= Female,
           SevTotal = Total) -> d1

df2 %>% select(3:5) %>% rename_with( ~ paste0("AttSv", .x)) -> d2

cbind(d1,d2) -> d3
d3 %>% mutate( pcAttSvMale   = 100 * AttSvMale/SevMale,
               pcAttSvFemale = 100 * AttSvFemale/SevFemale,
               pcAttSvTotal = 100 * AttSvTotal/SevTotal) -> d4

## Normal
TsSex("Cs14_NoDsbPriAge") -> n1
TsSex("Cs14NoDsbPriAttSex") -> n2

n1 %>% select(3:5) %>% rename_with( ~ paste0("Nm", .x)) -> n3
n2 %>% select(3:5) %>% rename_with( ~ paste0("AttNm", .x)) -> n4

cbind(n1[1:2],n3,n4) -> n5
head(n5,3)


# Combine two data sets
cbind(d4,n5[3:8]) -> PriAtt

rename(PriAtt,AreaCode = `Area code`)-> p2
inner_join(Loc2,p2, by = "AreaCode") -> p3

write.csv(p3,"p3.csv")

# Calculation
p3 %>% group_by(SR) %>%
    summarise(SevMale = sum(SevMale,na.rm=TRUE),
    SevFemale = sum(SevFemale,na.rm=TRUE),
    SevTotal = sum(SevTotal,na.rm=TRUE),
    AttSvMale = sum(AttSvMale,na.rm=TRUE),
    AttSvFemale= sum(AttSvFemale,na.rm=TRUE),
    AttSvTotal= sum(AttSvTotal,na.rm=TRUE),
    NmMale = sum(NmMale,na.rm=TRUE),
    NmFemale = sum(NmFemale,na.rm=TRUE),
    NmTotal = sum(NmTotal,na.rm=TRUE),
    AttNmMale = sum(AttNmMale,na.rm=TRUE),
    AttNmFemale = sum(AttNmFemale,na.rm=TRUE),
    AttNmTotal = sum(AttNmTotal,na.rm=TRUE)) -> SRPri
```
# School attendance rates of primary school age children 
## Tables

### Summary statistics
```{r sum, echo=FALSE}
SRPri %>% mutate(
            pcSevMaleAtt = AttSvMale * 100/SevMale,
            pcSevFemaleAtt = AttSvFemale * 100/SevFemale,
            pcSevTotalAtt = AttSvTotal * 100/SevTotal,
            pcNmMaleAtt = AttNmMale * 100/NmMale,
            pcNmFemaleAtt = AttNmFemale * 100/NmFemale,
            pcNmTotalAtt = AttNmTotal * 100/NmTotal) -> pcSRPri

s2 <- pcSRPri[c(1,16,19)]
names(s2) <-c("State/Region", "Severe", "Normal")
summary(s2)

```

### Attendance rates by State/region 
```{r table, echo=FALSE}
# Summarize state/region aggregates

print(s2)
```

## Visuals
### Boxplot
```{r bp, echo=FALSE, message=FALSE, results='hide',warning=FALSE}
melt(s2, id="State/Region") -> s3
```

```{r bp2, echo=FALSE}
ggplot(s3, aes(y=value, fill=variable)) + 
    geom_boxplot(width=0.5) +
    labs(fill="Disability status") +
    scale_x_discrete(labels = NULL, breaks = NULL) +
    scale_fill_manual(name="",
                      values=c("#9999FF","#66FFFF"),
                      labels=c("Severe","Normal")) +
    ylab("percent") + ylim(0,100) +
    theme_minimal() +
    theme(legend.position = c(0.2, 0.5))
```

### Group bar

```{r gb2, echo=FALSE}
ggplot(s3, aes(y=value, x=`State/Region`, fill = variable)) +
    geom_bar(position="dodge", stat="identity") +
    scale_fill_manual(name="",
                      values=c("#7F00FF","#33FF33" ),
                      labels=c("Severe","Normal"))  +
    xlab ("State and Region") + ylab ("Percent") +
    theme_minimal()+
    theme(legend.position = "top") +
    theme(axis.text.x = element_text(angle=90)) +
    ggtitle("Primary School Attendance")
```

```{r dt, echo=FALSE}
names(s3) <- c("State/Region","Disability Status", "Attendance(%)") 
options(digits = 3)
kable(s3, caption= "<Heading 1> Primary School Attendance rates",format ="html") %>%  
    footnote(general = "", general_title = "N.B. The NA values are on the children with disaiblity side that might also mean their being neglected status.")
```

# Conclusion
The analytical results from Myanmar Census 2014 show that the children having *severe disability* were **hugely** lagging behind in their primary school attendance compared to those of the children with *no disability*.

# Appendix

## Township attendance rates
```{r print_tb, echo=FALSE, results= 'asis', warning=FALSE, message=FALSE}
p3 %>% select(1,4,contains("pc")) -> p4
names(p4) <-c("State/Region","Census Township", "Male Severe","Female Severe","Total Severe","Male Normal","Female Normal","Total Normal")
```
```{r print_tb2, echo=FALSE, results= 'asis', warning=FALSE, message=FALSE}
options(digits = 3)
kable(p4, caption= "<Heading 1> Primary School Attendance rates",format ="html") %>%  
    footnote(general = "", general_title = "N.B. The NA values are on the children with disaiblity side that might also mean their being neglected status.")
```
